Scattering Auditory Scenes for Classification: Supervised and Unsupervised Approaches

### 2016-06-27 paper/structureScene.tex
write more of section IV

despite its size, it is still challenging as state-of-the-art systems achieves 76 \% \cite{6701890} (winner of the 2013 challenge RQA+SVM). For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively. On the latter, the state-of-the-art is currently 95 \% \cite{bisot2016acoustic}.

@@

In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach.
However, it favors the center of the scene while attenuating the presence of lateral sources.
Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
Instead, so as to take advantage of stereophonic data, we perform multiiple combinations of the left channel $\boldsymbol{x_{\mathsf{L}}}(t)$ and right channel $\boldsymbol{x_{\mathsf{R}}}(t)$ into monophonic signals $\boldsymbol{x_{\alpha}}(t)$ according to the equation
\begin{equation}
\boldsymbol{x_{\alpha}}(t) =
\dfrac{1+\alpha}{2} \boldsymbol{x_\mathsf{L}}(t) +
\dfrac{1-\alpha}{2} \boldsymbol{x_\mathsf{R}}(t),
\end{equation}
where $\alpha$ is a panoramic mixing weight, taking values between $-1$ and $+1$, which could be deterministic or random.
In following experiments, we deterministically set $\alpha$ to the values $-1$, $\frac{1}{2}$, $0$, $+\frac{1}{2}$, and $+1$.
This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.

\subsection{Feature selection}
The starting scale $\gamma_{1,\min}$ corresponding to some upper bound $\xi_{1,\max}$ on the frequency axis is given by
\begin{equation}
\gamma_{1,\min} =
\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil
\end{equation}
where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.

\subsection{Logarithmic compression}
Many pattern recognition algorithms, including nearest neighbour classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.

%Logarithmic compression is corroborated by psychoacoustics
% linearizes the dependency with respect to the distance to the source

@@

@@

\subsection{Early vs. late integration}
Fine-grained classification of acoustic scenes is only made
possible by integrating signal information over a long temporal context.

\cite{Joder2009}

The next section discusses more advanced methods for the temporal integration of scattering coefficients in an unsupervised setting.

>>>>

despite its size, it is still challenging as state-of-the-art systems achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM). For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively. On the latter, the state-of-the-art is currently 95 \% \cite{bisot2016acoustic}.

@@

In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach.
However, it favors the center of the scene while attenuating the presence of lateral sources.
Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
Instead, so as to take advantage of stereophonic data, we perform multiiple combinations of the left channel $\boldsymbol{x_{\mathsf{L}}}(t)$ and right channel $\boldsymbol{x_{\mathsf{R}}}(t)$ into monophonic signals $\boldsymbol{x_{\alpha}}(t)$ according to the equation
\begin{equation}
\boldsymbol{x_{\alpha}}(t) =
\dfrac{1+\alpha}{2} \boldsymbol{x_\mathsf{L}}(t) +
\dfrac{1-\alpha}{2} \boldsymbol{x_\mathsf{R}}(t),
\end{equation}
where $\alpha$ is a panoramic mixing weight, taking values between $-1$ and $+1$, which could be deterministic or random.

In following experiments, we deterministically set $\alpha$ to the values $-1$, $\frac{1}{2}$, $0$, $+\frac{1}{2}$, and $+1$.
This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.
At test time, only the center mix ($\alpha = 0$) is used to perform a prediction.

\subsection{Feature selection}
In a scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
This has lead \cite{roma2013}

The starting scale $\gamma_{1,\min}$ corresponding to some upper bound $\xi_{1,\max}$ on the frequency axis is given by
\begin{equation}
\gamma_{1,\min} =
\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil
\end{equation}
where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.

\subsection{Logarithmic compression}
Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design, and their distribution is typically skewed towards the right.

This issue can be addressed by applying to all coefficients a pointwise concave transformation, \eg logarithmic, hence reducing skewness.
In particular, logarithmic compression is ubiquitous in audio signal processing, as it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
In the context of auditory scenes, we must also stress that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone.
Logarithmic compression can linearize this dependency, which arguably facilitates the construction of an invariant to this distance at the classifier stage.

@@

\subsection{Early vs. late integration}
Fine-grained classification of acoustic scenes is only made
possible by integrating signal information over a long temporal context.
A straightforward approach, known as early integration, consists in summarizing the multivariate time series of scattering coefficients

Early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff. We refer to \cite{Joder2009} for a comprehensive review of this problematic.

@@

The next section discusses adaptive methods for the temporal integration of scattering coefficients in an unsupervised setting.
